---
title: "Guanyingdong Stone Artefact Assemblage Report"
author: "HY and BM"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   bookdown::word_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
# read in the data
library(plyr)
library(dplyr)
library(knitr)
library(readxl)
# install.packages("readxl")
file_name <- "../data/artefacts (27 Jul 2016).xls" 
flakes <- read_excel(file_name, sheet = "flake basics")
cores <- read_excel(file_name, sheet = "core basics")
debris <- read_excel(file_name, sheet = "chunk&debris")
retouch <- read_excel(file_name, sheet = "retouch")
```

```{r}
# are the artefact numbers really unique?
# sum(duplicated(flakes$number))
# extract duplicated flakes 
flakes_duplicated <- flakes[ duplicated(flakes$number), ]

# remove from original data
flakes <- flakes[!duplicated(flakes$number), ]

# update the artefact ID so it's not duplicate
flakes_duplicated$number <- paste0(flakes_duplicated$number, "a")

# put back into original data 
flakes <- rbind(flakes, flakes_duplicated)

# do the same for the retouch...

# are the arteact numbers really unique?
# sum(duplicated(retouch$number))

# extract duplicated retouch 
retouch_duplicated <- retouch[ duplicated(retouch$number), ]

# remove from original data
retouch <- retouch[!duplicated(retouch$number), ]

# update the artefact ID so it's not duplicate
retouch_duplicated$number <- paste0(retouch_duplicated$number, "a")

# put back into original data 
retouch <- rbind(retouch, retouch_duplicated)

############
# basic details of the dataset
artefact_ids <- list(flake_ids = flakes$number,
                     core_ids = cores$number,
                     debris_ids = debris$number,
                     retouch_ids = retouch$number)



# how many unique artefacts?
totals_of_each_type <- lapply(artefact_ids, function(i) length(unique(i)))
count_unique_artefacts <- sum(unlist(totals_of_each_type))
# debris pieces that are  retouch pieces
debris_with_retouch <- intersect(artefact_ids$retouch_ids, artefact_ids$debris_ids)
proportion_debris_retouched <- length(debris_with_retouch) / count_unique_artefacts 
# proportion of each type
library(knitr)
type_table <- data.frame(type = names(as.data.frame(totals_of_each_type)),
                         count = as.numeric(as.data.frame(totals_of_each_type)),
                         proportion = round(as.numeric(prop.table(as.data.frame(totals_of_each_type))),3))
# add total row at the bottom
type_table$type <- as.character(type_table$type)
type_table <- rbind(type_table, c("total", count_unique_artefacts, round(sum(type_table$proportion),2)))
# print pretty table                         
# kable(type_table)
```


# Introduction 

The Guanyindong site, located in  Guanyindong village, Qianxi County of Guizhou Province (26°51′26″N, 105°58′7″E) at an elevation of 1464 m a.s.l., is a limestone cave site extending from east to west it was discovered by a team organized by the institute of Vertebrate Paleontology and Paleoanthropolgy(IVPP),Chinese Academy of Sciences in 1964. Several excavations were conducted in 1965, 1972 and 1973, yeilding   A total of `r type_table$count[2]` cores, `r type_table$count[1]` flakes, `r type_table$count[4]` retouched pieces and `r type_table$count[3]` pieces of debris were identified. 

Introduction of paleolithc research in south Asia (or China). 

Introduction the distribution of levollois technique (origin, dispersion, distribution). 

Prolem: East Asia, why people thought no levollois. Why studying this site is important. Aim in this study.

## Raw materials

```{r rawmaterials}
flake_raw_material <- table(flakes$material)

#' Data frame of counts, frequences and proportion of values in a vector
#' 
#' @param x a vector

tbl_of_counts_and_props <- function(x){
  tbl <- table(x)
  res <- data.frame(tbl, 
                    round(prop.table(tbl) * 100, 2))
  colnames(res) <- c('Name', 'Count', 'Percentage', 'Proportion')
  res
}

raw_material_tbl <- tbl_of_counts_and_props(flakes$material)

```

The assemblage is dominated by chert (`r raw_material_tbl$Proportion[2]`%) followed by limestone (`r raw_material_tbl$Proportion[3]`%) and small amounts of basalt, sandstone and quartz. 


## Flakes

```{r flakebasics}
flake_length_mean <- round(mean(flakes$mass, na.rm = TRUE),1) #there are thousands of flakes

```


The average maximum length of the flakes is `r flake_length_mean` mm

## Cores

## Retouched pieces

### Indices

## Levallois 

# Discussion

1) Comparing with other sites in south China, Guanyindong is featured by the appearance of Levallois technique.Why do you think Gyd has levallois? Previous results from other sites in South China suggest that they have no levallois techinique (give examples and discussions)

2) In the lithic assemblage, we found 59 Levallois flakes, 11 Levallois cores, with distinguishable characters from Europe and Africa(less proportion, relative more proto morphology). What are the main difference between Gyd and other sites. What is the implication of such difference. Are there any difference in the Levallois techniques in Europe and Africa or other Asia sites?

3) Provide detailed discussion on the Demographic model, and explain the reason why Levallois technique in Southeast Asia is different from western. China and Southeast Asia is geographically distant from East Africa where Levallois technique is originated, as hominids dispersal from western to eastern, progressively smaller population, drastic changes of environment have made an inevitable influence on the style of Levallois technique. 

4) Drawbacks (assumptions) of the model. Other possible explanation you can offer.



# Conclusion
