---
title: "Guanyingdong Stone Artefact Assemblage Report: Compare artefacts from upper and lower parts of the deposit"
author: "HY and .BM"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   bookdown::word_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
---

```{r setup-flake-size, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Nature’s standard figure sizes are 89 mm (single column) and 183 mm (double column)
 # https://www.nature.com/nature/for-authors/formatting-guide
```

Get the basic flake details, run from L454 of hu_artefacts_report.Rmd, then, we are checking for differences between  upper and lower

# Counts

Altogether:

```{r}
flakes %>% 
  select(layer) %>% 
  bind_rows(cores %>% 
              select(layer)) %>% 
  group_by(layer) %>% 
  tally() %>% 
  mutate(perc = n / sum(n) * 100)
```

Flakes:

```{r}
flakes %>% 
  group_by(layer) %>% 
  tally() %>% 
  mutate(perc = n / sum(n) * 100)
```

Cores:

```{r}
cores %>% 
  group_by(layer) %>% 
  tally() %>% 
  mutate(perc = n / sum(n) * 100)
```

In the main text we write: "Among these, 5 were identified as Levallois, and 3 of them (including 2 cores and 1 flake) are from Group A and 2 (1 flake and 1 tool) are from Group B."

upper - Group A: 2 cores and 1 flake
lower - Group B: 0 cores and 1 flake and 1 tool

HY says the Levallois artefacts are coded LVF in the type1 column. But we don't have any that are coded with upper/lower in the layer column. So I need some advice about exactly which they are. 

HY says:

P5348 is from the upper [but not labelled in the spreadsheet]
P5306 is from the lower, is a levallois tool, because it is retouched
P5366 is from the lower, is a levallois flake


```{r}
flakes %>% 
  filter(layer %in% c('lower', 'upper')) %>% 
  group_by(layer, type1) %>% 
  tally()
```

# Technological and metric comparisons

Flake metrics:

```{r}
flakes %>% 
  filter(!is.na(layer)) %>% 
  select(which(sapply(., class) == "numeric"), layer) %>% 
  filter(cortex_percentage < 20) %>% 
  gather(variable, value, -layer) %>% 
  mutate(variable = str_to_title(str_replace_all(variable, "_", " "))) %>% 
  filter(!is.na(value)) %>% 
  mutate(layer = factor(layer)) %>% 
  ggplot(aes(layer,
             value)) +
  geom_boxplot() +
  facet_wrap( ~ variable, scales = "free_y") +
  xlab("Size class") +
  ylab("") +
  theme_minimal()
```

Flake technological attributes



```{r}

upper_lower_flakes_categorical <- 
flakes %>% 
  select_if(function(col) is.character(col) | 
                                   all(col == .$layer)) %>% 
  select(-one_of(c("bag",  "Li_classifi", "note", "number"))) %>% 
  gather(variable, value, -layer) %>% 
  mutate(layer = factor(layer)) %>% 
  filter(!is.na(value)) %>% 
  group_by(layer, variable, value) %>% 
  tally() %>% 
  ungroup() %>% 
  group_by(layer, variable) %>%  
  mutate(prop = n / sum(n) * 100) 


# tidy up some of these variables
library(stringr)
library(forcats)

upper_lower_flakes_cats <- 
flakes %>% 
  select_if(function(col) is.character(col) | 
                                   all(col == .$layer)) %>% 
   select(-one_of(c("bag", "Li_classifi", "note", "number"))) %>% 
  filter(!is.na(layer))


# raw material -----------------------------------------------------------------
# material, ok
upper_lower_flakes_cats %>% 
 group_by(material) %>% 
 tally(sort = TRUE)

# make table for plotting
upper_lower_flakes_cats_raw_material <- 
upper_lower_flakes_cats %>% 
rename(`Raw material` = material) %>% 
 group_by(`Raw material` , layer) %>% 
 tally(sort = TRUE)

# flake type -----------------------------------------------------------------

# type1, ok
upper_lower_flakes_cats %>% 
 group_by(type1) %>% 
 tally(sort = TRUE)

upper_lower_flakes_cats_flake_types <- 
  upper_lower_flakes_cats %>% 
  mutate(type1 = if_else(str_detect(type1, "lvf"), "Levallois", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "ret"), "retouch", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "blade|bld"), "blade", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "flk|flake"), "flake", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "dbd"), "débordant", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "kbw"), "Kombewa", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "end"), "end scraper", type1)) %>% 
  mutate(type1 = fct_lump(type1, 2))

# check again
upper_lower_flakes_cats_flake_types %>% 
 group_by(type1) %>% 
 tally(sort = TRUE)

# for plotting
flakes_clustered_cats_flake_types <- 
flakes_clustered_cats_flake_types %>% 
 rename(`Flake type` = type1) %>% 
 group_by(`Flake type`, cluster) %>% 
 tally(sort = TRUE)

# retouch type -----------------------------------------------------------------

# type 2, special case , retouch types
flakes_clustered_cats %>% 
 group_by(type2) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_retouch_type <- 
flakes_clustered_cats %>% 
 separate(type2, into = c("type2a", "type2b", "type2c"), sep = ",") %>% 
  mutate(type2a = if_else(str_detect(type2a, "lvf"), "levallois", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "ret"), "retouch", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "blade|bld"), "blade", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "flk|flake"), "flake", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "tang"), "tanged", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "not"), "notch", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "bur"), "burin", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "point"), "point", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "dent"), "denticulate", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "scp"), "scraper", type2a)) %>% 
  mutate(type2a = str_squish(type2a))  %>% 
  mutate(type2a = fct_lump(type2a, 7)) %>% 
  filter(!is.na(type2a)) 

# check it
flakes_clustered_cats_retouch_type %>% 
 group_by(type2a) %>% 
 tally(sort = TRUE)

# for plotting
flakes_clustered_cats_retouch_type <- 
flakes_clustered_cats_retouch_type %>% 
 rename(`Retouch type` = type2a ) %>% 
 group_by(`Retouch type`, cluster) %>% 
 tally(sort = TRUE)
  

# platform type -----------------------------------------------------------------

# Platform, ok
flakes_clustered_cats %>% 
 group_by(platform) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_platform_type  <- 
  flakes_clustered_cats %>% 
   mutate(platform = if_else(str_detect(platform, "facet"), "facetted", platform)) %>% 
   mutate(platform = if_else(str_detect(platform, "cotex|cortex"), "cortex", platform)) %>%
   mutate(platform = fct_lump(platform, 6)) %>% 
  filter(!is.na(platform))


# for plotting
flakes_clustered_cats_platform_type <- 
flakes_clustered_cats_platform_type %>% 
 rename(`Platform type` = platform ) %>% 
 group_by(`Platform type`, cluster) %>% 
 tally(sort = TRUE)
  

# platform shape -----------------------------------------------------------------

# platform shape, ok
flakes_clustered_cats %>% 
 group_by(platform_shape) %>% 
 tally(sort = TRUE)


flakes_clustered_cats_platform_shape <- 
flakes_clustered_cats %>% 
  mutate(platform_shape = if_else(platform_shape == "sb", "quadrangle", platform_shape)) %>%
  mutate(platform_shape = if_else(platform_shape == "cdg", "chapeau de gendarme", platform_shape)) %>%
  mutate(platform_shape = if_else(platform_shape == "sx", "rhombus", platform_shape)) %>%
  mutate(platform_shape = fct_lump(platform_shape, 4))  %>% 
  filter(!is.na(platform_shape))

# for plotting
flakes_clustered_cats_platform_shape <- 
flakes_clustered_cats_platform_shape %>% 
 rename(`Platform shape` = platform_shape ) %>% 
 group_by(`Platform shape`, cluster) %>% 
 tally(sort = TRUE)
 
# cortex texture-----------------------------------------------------------------
 
# cortex texture, ok 
flakes_clustered_cats %>% 
 group_by(cortex_texture) %>% 
 tally(sort = TRUE)

# for plotting
flakes_clustered_cats_cortex_texture <- 
flakes_clustered_cats %>% 
 rename(`Cortex texture` = cortex_texture ) %>% 
 group_by(`Cortex texture`, cluster) %>% 
 filter(!is.na(`Cortex texture` )) %>% 
 tally(sort = TRUE)

# cortex location-----------------------------------------------------------------
 
# cortex location, ok
flakes_clustered_cats %>% 
 group_by(cortex_location) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_cortex_location <- 
  flakes_clustered_cats %>% 
   mutate(cortex_location = if_else(str_detect(cortex_location, "prox"), 
                                    "proximal", 
                                    cortex_location)) %>%
   mutate(cortex_location = fct_lump(cortex_location, 6)) %>% 
  filter(!is.na(cortex_location),
         cortex_location != "vent")

flakes_clustered_cats_cortex_location <- 
flakes_clustered_cats_cortex_location %>% 
 rename(`Cortex location` = cortex_location ) %>% 
 group_by(`Cortex location`, cluster) %>% 
 filter(!is.na(`Cortex location` )) %>% 
 tally(sort = TRUE)

# cortex location-----------------------------------------------------------------

# initiation, ok 
flakes_clustered_cats %>% 
 group_by(initiation) %>% 
 tally(sort = TRUE)

flakes_clustered_cats_initiation <- 
  flakes_clustered_cats %>% 
   mutate(initiation = fct_lump(initiation, 3))

flakes_clustered_cats_initiation <- 
flakes_clustered_cats_initiation %>% 
 rename(`Initiation` = initiation ) %>% 
 group_by(`Initiation`, cluster) %>% 
 filter(!is.na(`Initiation` )) %>% 
 tally(sort = TRUE)

# cortex location-----------------------------------------------------------------

# platform preparation, ok, ignore, no meaningful info in here
flakes_clustered_cats %>% 
 group_by(platform_preparation) %>% 
 tally(sort = TRUE)

flakes_clustered_cats <- 
  flakes_clustered_cats %>% 
   mutate(platform_preparation = if_else(str_detect(platform_preparation, "fac"), "facetted", platform_preparation)) %>% 
   mutate(platform_preparation = fct_lump(platform_preparation, 3))

```

Try for a nice plot, we have separate dataframes for each variable. 

```{r}
list_of_dfs <- list(flakes_clustered_cats_retouch_type,
                    flakes_clustered_cats_raw_material,
                    flakes_clustered_cats_platform_type,
                    flakes_clustered_cats_platform_shape,
                    flakes_clustered_cats_initiation,
                    flakes_clustered_cats_flake_types,
                    flakes_clustered_cats_cortex_texture,
                    flakes_clustered_cats_cortex_location)

map_dfr(list_of_dfs, ~nest(.x))

all_the_categorical_variables <- 
  map_dfr(list_of_dfs, 
          ~gather(.x, variable, value, -cluster, -n))

# Compute proportions per category

flakes_clustered_cats_long <- 
all_the_categorical_variables %>% 
  mutate(value = str_squish(value)) %>% 
  mutate(cluster = factor(cluster)) %>% 
  group_by(cluster, variable) %>%  
  mutate(prop = n / sum(n) * 100)  %>% 
  arrange(cluster, variable)
```


Here is the plot

```{r}
categorical_plots <- 
flakes_clustered_cats_long %>% 
  ungroup(cluster) %>% 
  nest(-variable) %>% 
  mutate(data = map(data, ~.x %>% mutate(value = as.character(value)))) %>% 
  mutate(plot = map2(data,
                     variable, ~.x %>%  
                            ggplot(aes(cluster, 
                                          prop, 
                                          fill = value)) + 
                       scale_fill_viridis_d() +
                            geom_col() +
                            xlab("") +
                            ylab("%") +
                       theme_minimal(base_size = 8) +
                           theme(legend.key.size = unit(0.4,"line"),
                                 legend.key.width = unit(0.4,"line"),
                                 legend.key.height = unit(0.4,"line"),
                                 legend.text = element_text(size=8), 
                                 legend.title = element_blank(),
                                 legend.margin=margin(t=-0.5, 
                                                      r=-0.5, 
                                                      b=-0.5, 
                                                      l=-0.1, 
                                                      unit="cm"),
                                 legend.box.margin=margin(-0.1,
                                                          -0.1,
                                                          -0.1,
                                                          -0.3,
                                                          unit="cm"),
                                 plot.margin = unit(x = c(0.5, 
                                                          0, 
                                                          0,
                                                          0.5), 
                                                    units = "cm")) +
                            ggtitle(str_glue('{.y}'))))
   

# arrange and draw the plots
categorical_plots_arranged <- 
cowplot::plot_grid(plotlist = categorical_plots$plot,
                   align = "hv",
                   axis = "tblr",
                   ncol = 2)

save_plot("analysis/report/figure_two.png", categorical_plots_arranged,
          base_aspect_ratio = 1,
          base_height = 8,
          base_width = 18.3/2.5
)

```


Compare cores from upper and lower

