---
title: "Guanyingdong Stone Artefact Assemblage Report: Variation by Size"
author: "HY and BM"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   bookdown::word_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
---

```{r setup-flake-size, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

THe reviewers ask:

>  can they offer other data that might elucidate changes in flaking type with reduction, e.g. changes in flake attributes that correlate with size declines?


Read in the data, deal with dupicates, get the basic flake details, run from L454 of hu_artefacts_report.Rmd

Let's see if there are any natural groupings in flake mass:

```{r}

flake_hist_1 <- 
ggplot(flakes,
       aes(mass)) +
  geom_histogram() +
  theme_minimal() +
  xlab("Mass (g)")

flake_hist_2 <- 
ggplot(flakes,
       aes(mass)) +
  geom_histogram() +
  theme_minimal() +
  scale_x_log10() +
  xlab("Mass (g)")


library(cowplot)
flake_mass_distributions <- plot_grid(flake_hist_1, flake_hist_2, ncol = 1)
```

Nothing obvious, let's try a little clustering. We will try univariate clustering:

```{r}
require(Ckmeans.1d.dp)
result <- Ckmeans.1d.dp(flakes$mass[!is.na(flakes$mass)])
plot(result)

n_clusters <- length(result$size)
n_clusters_30 <- length(result$size[result$size > 30])

# assign cluster ID to table
flakes_clustered <- 
flakes %>% 
  filter(!is.na(mass)) %>% 
  mutate(cluster = result$cluster) %>% 
  right_join(flakes) %>% 
  filter(cluster %in% 1:n_clusters_30)
```


We get `r n_clusters` clusters, with `r n_clusters_30` that have more than 30 artefacts in them

```{r}
ggplot(flakes_clustered,
       aes(mass, 
          length,
          colour = factor(cluster)))+
  geom_point()

flake_hist_with_clusters <- 
ggplot(flakes_clustered,
       aes(mass)) +
  geom_histogram(aes(fill = factor(cluster))) +
  scale_fill_viridis_d(name = "Size class") +
  theme_minimal() +
  xlab("Mass (g)") 



```

Now let's see how some of the key variables change accross these size clusters:

```{r}
library(stringr)
flakes_clustered_numeric_vars_plot <- 
flakes_clustered %>% 
  select_if(is.numeric) %>% 
  gather(variable, value, -cluster) %>% 
  filter(!is.na(value)) %>% 
  mutate(cluster = factor(cluster)) %>% 
  mutate(variable = str_replace_all(variable, "_", " ")) %>% 
  ggplot(aes(cluster,
             value)) +
  geom_boxplot(position = position_dodge2(preserve = "total")) +
  facet_wrap( ~ variable, scales = "free_y") +
  theme_minimal(base_size = 8)
```

We see that some dimensions don't increase with size. Platform thickness decreases a little, as mass goes up, platform width becomes more variable. Scar numbers have less high outliers as size increases. 

What if we just look at flakes with very little cortex, and thus exclude the primary reduction flakes:

```{r}
flakes_clustered_numeric_vars_low_cortex_plot <- 
flakes_clustered %>% 
  select_if(is.numeric) %>% 
  filter(cortex_percentage < 20) %>% 
  gather(variable, value, -cluster) %>% 
  filter(!is.na(value)) %>% 
  mutate(cluster = factor(cluster)) %>% 
  ggplot(aes(cluster,
             value)) +
  geom_boxplot() +
  facet_wrap( ~ variable, scales = "free_y") +
  theme_minimal()
```

Platform thickness decreases, width at 50% and 75% does not increase as much as width at 25% suggesting a slight increase in triangular-shaped flakes over parallel-sided flakes as the flakes get larger. 

How about the categorical variables?

```{r}

flakes_categorical <- 
flakes_clustered %>% 
  filter(cortex_percentage < 20) %>% 
  select_if(function(col) is.character(col) | 
                                   all(col == .$cluster)) %>% 
  select(-one_of(c("bag", "layer", "Li_classifi", "note", "number"))) %>% 
  gather(variable, value, -cluster) %>% 
  mutate(cluster = factor(cluster)) %>% 
  filter(!is.na(value)) %>% 
  group_by(cluster, variable, value) %>% 
  tally() %>% 
  ungroup() %>% 
  group_by(cluster, variable) %>%  
  mutate(prop = n / sum(n) * 100) 


# tidy up some of these variables
library(stringr)
library(forcats)

flakes_clustered_cats <- 
flakes_clustered %>% 
  select_if(function(col) is.character(col) | 
                                   all(col == .$cluster)) %>% 
   select(-one_of(c("bag", "layer", "Li_classifi", "note", "number")))

# material, ok
flakes_clustered_cats %>% 
 group_by(material) %>% 
 tally(sort = TRUE)

# type1, ok
flakes_clustered_cats %>% 
 group_by(type1) %>% 
 tally(sort = TRUE)

flakes_clustered_cats <- 
  flakes_clustered_cats %>% 
  mutate(type1 = if_else(str_detect(type1, "lv|lev"), "Levallois", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "ret"), "retouch", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "blade|bld"), "blade", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "flk|flake"), "flake", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "dbd"), "débordant", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "kbw"), "Kombewa", type1)) %>% 
  mutate(type1 = if_else(str_detect(type1, "end"), "end scraper", type1)) %>% 
  mutate(type1 = fct_lump(type1, 5))

# type 2, special case , retouch types
flakes_clustered_cats %>% 
 group_by(type2) %>% 
 tally(sort = TRUE)

flakes_clustered_cats <- 
flakes_clustered_cats %>% 
 separate(type2, into = c("type2a", "type2b", "type2c"), sep = ",") %>% 
  mutate(type2a = if_else(str_detect(type2a, "lv|lev"), "levallois", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "ret"), "retouch", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "blade|bld"), "blade", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "flk|flake"), "flake", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "tang"), "tanged", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "not"), "notch", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "bur"), "burin", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "point"), "point", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "dent"), "denticulate", type2a)) %>% 
  mutate(type2a = if_else(str_detect(type2a, "scp"), "scraper", type2a)) %>% 
  mutate(type2a = str_squish(type2a))  %>% 
  mutate(type2a = fct_lump(type2a, 10)) %>% 
  filter(!is.na(type2a))

# check it
flakes_clustered_cats %>% 
 group_by(type2a) %>% 
 tally(sort = TRUE)
  
# Platform, ok
flakes_clustered_cats %>% 
 group_by(platform) %>% 
 tally(sort = TRUE)

flakes_clustered_cats <- 
  flakes_clustered_cats %>% 
   mutate(platform = if_else(str_detect(platform, "facet"), "facetted", platform)) %>% 
   mutate(platform = if_else(str_detect(platform, "cotex|cortex"), "cortex", platform)) %>%
   mutate(platform = fct_lump(platform, 6)) %>% 
  filter(!is.na(platform))
  
# platform shape, ok
flakes_clustered_cats %>% 
 group_by(platform_shape) %>% 
 tally(sort = TRUE)

n <- 10
platform_shape_top <- 
flakes_clustered_cats %>% 
 group_by(platform_shape) %>% 
 tally(sort = TRUE) %>% 
filter(n >= 10)

flakes_clustered_cats <- 
flakes_clustered_cats %>% 
  mutate(platform_shape = if_else(platform_shape == "sb", "quadrangle", platform_shape)) %>%
  mutate(platform_shape = if_else(platform_shape == "cdg", "chapeau de gendarme", platform_shape)) %>%
  mutate(platform_shape = if_else(platform_shape == "sx", "rhombus", platform_shape)) %>%
  mutate(platform_shape = fct_lump(platform_shape, 4))  %>% 
  filter(!is.na(platform_shape))

# cortex texture, ok -- up to here --
flakes_clustered_cats %>% 
 group_by(cortex_texture) %>% 
 tally(sort = TRUE)

# cortex location, ok
flakes_clustered_cats %>% 
 group_by(cortex_location) %>% 
 tally(sort = TRUE)

flakes_clustered_cats <- 
  flakes_clustered_cats %>% 
   mutate(cortex_location = if_else(str_detect(cortex_location, "prox"), 
                                    "proximal", 
                                    cortex_location)) %>%
   mutate(cortex_location = fct_lump(cortex_location, 6)) %>% 
  filter(!is.na(cortex_location),
         cortex_location != "vent")

# initiation, ok 
flakes_clustered_cats %>% 
 group_by(initiation) %>% 
 tally(sort = TRUE)

flakes_clustered_cats <- 
  flakes_clustered_cats %>% 
   mutate(initiation = fct_lump(initiation, 3))

# platform preparation, ok, ignore, no meaningful info in here
flakes_clustered_cats %>% 
 group_by(platform_preparation) %>% 
 tally(sort = TRUE)

flakes_clustered_cats <- 
  flakes_clustered_cats %>% 
   mutate(platform_preparation = if_else(str_detect(platform_preparation, "fac"), "facetted", platform_preparation)) %>% 
   mutate(platform_preparation = fct_lump(platform_preparation, 3))

```

Try for a nice plot

```{r}
flakes_clustered_cats_long <- 
flakes_clustered_cats %>% 
  gather(variable, value, -cluster)  %>% 
  filter(!is.na(value), 
         variable != "comment",
         variable != "heat_damage",
         variable != "platform_preparation",
         variable != "type2c",
         variable != "type2b",
         variable != "box") %>% 
  mutate(variable = case_when(
    variable == 'type1' ~ "Flake type",
    variable == 'type2a' ~ "Retouch type",
    variable == 'cortex_location' ~ "Cortex location",
    variable == 'cortex_texture' ~ "Cortex texture",
    variable == 'initiation' ~ "Initiation",
    variable == 'material' ~ "Raw material",
    variable == 'platform' ~ "Platform type",
    variable == 'platform_shape' ~ "Platform shape",
    TRUE ~ variable)) %>% 
  mutate(value = str_squish(value)) %>% 
  mutate(cluster = factor(cluster)) %>% 
  group_by(cluster, variable, value) %>% 
  tally() %>% 
  ungroup() %>% 
  group_by(cluster, variable) %>%  
  mutate(prop = n / sum(n) * 100) 
```


Here is the plot

```{r}
categorical_plots <- 
flakes_clustered_cats_long %>% 
  ungroup(cluster) %>% 
  nest(-variable) %>% 
  mutate(plot = map2(data,
                     variable, ~.x %>%  
                            ggplot(aes(cluster, 
                                          prop, 
                                          group=1, 
                                          fill = value)) + 
                       scale_fill_viridis_d() +
                            geom_col() +
                            xlab("") +
                            ylab("%") +
                           theme(legend.key.size = unit(0.4,"line"),
                                 legend.key.width = unit(0.4,"line"),
                                 legend.key.height = unit(0.4,"line"),
                                 legend.text = element_text(size=8), 
                                 legend.title = element_blank()) +
                            ggtitle(str_glue('{.y}'))))
   

# arrange and draw the plots
categorical_plots_arranged <- 
cowplot::plot_grid(plotlist = categorical_plots$plot,
                   align = "hv",
                   axis = "tblr")

```




Try this on the Lev flakes only:

```{r}
flakes_clustered %>% 
  select_if(function(col) is.numeric(col) | 
                                   all(col == .$type1))  %>% 
  filter(str_detect(type1, "LV|leva")) %>% 
    select(-box, -type1) %>% 
  gather(variable, value, -cluster) %>% 
  filter(!is.na(value)) %>% 
  mutate(cluster = factor(cluster)) %>% 
  ggplot(aes(cluster,
             value)) +
  geom_boxplot() +
  facet_wrap( ~ variable, scales = "free_y") +
  theme_minimal()
```


```{r}
flakes_clustered_cats_long_leva <- 
flakes_clustered_cats %>% 
  filter(type1 %in% c("levallois", "LVF")) %>% 
  gather(variable, value, -cluster)  %>% 
  filter(!is.na(value), 
         variable != "comment",
         variable != "heat_damage",
         variable != "platform_preparation",
         variable != "type2c") %>% 
  mutate(value = str_squish(value)) %>% 
  mutate(cluster = factor(cluster)) %>% 
  group_by(cluster, variable, value) %>% 
  tally() %>% 
  ungroup() %>% 
  group_by(cluster, variable) %>%  
  mutate(prop = n / sum(n) * 100)  

categorical_plots_leva <- 
flakes_clustered_cats_long_leva %>% 
  ungroup(cluster) %>% 
  nest(-variable) %>% 
  mutate(plot = map(data , ~.x %>%  
                            ggplot(aes(cluster, 
                                          prop, 
                                          group=1, 
                                          fill = value)) + 
                            geom_col() +
                            xlab("") +
                           theme(legend.key.size = unit(0.4,"line"),
                                 legend.key.width = unit(0.2,"line"),
                                 legend.key.height = unit(0.4,"line"),
                                 legend.text = element_text(size=8)))) 

# get legend labels so we know what variable is plotted
# in each plot, not sure why I can't do this with map
for(i in 1:nrow(categorical_plots_leva)){
  categorical_plots_leva$plot[[i]] <- 
    categorical_plots_leva$plot[[i]] + guides(fill=guide_legend(title=unique(categorical_plots_leva$variable)[i]))
}

# arrange and draw the plots
cowplot::plot_grid(plotlist = categorical_plots_leva$plot,
                   align = "hv",
                   axis = "tblr")


```


